### 2022-01-17 manuscript/submission_old.tex
revisao geophysics inclusas

We present a fast equivalent layer method for processing large-scale magnetic data. 
We demonstrate that the sensitivity matrix associated with an equivalent layer
of dipoles can be arranged to a Block-Toeplitz Toeplitz-Block (BTTB) structure for the 
case in which observations and dipoles are aligned on a 
horizontal and regularly-spaced grid.
The product of a BTTB matrix and an arbitrary vector represents a discrete 
convolution and can be efficiently computed via 2D Fast Fourier Transform.
In this case, the matrix-vector product uses only the elements forming the first column
of the BTTB matrix, saving computational time and memory. 
Our convolutional equivalent layer method uses this approach to compute 
the matrix-vector products in the iterative conjugate gradient algorithm with the purpose 
of estimating the physical-property distribution over the equivalent layer for 
large data sets.
Synthetic tests with a mid-size $100 \times 50$ grid of total-field anomaly data
show a decrease of $\approx 10^4$ in floating-point operations and $\approx 25\times$ 
in computation runtime of our method compared to the classical approach of solving
the least-squares normal equations via Cholesky decomposition. 
Faster results are obtained for millions of data, showing drastic decreases in RAM usage
and runtime, allowing to perform magnetic data processing of large data sets on regular 
desktop computers. 
Our results also show that, compared to the classical Fourier approach, the magnetic
data processing with our method requires similar computation time, but produces significantly 
smaller border effects without using any padding scheme and also is more robust to 
deal with data on irregularly spaced points or on undulating observation surfaces.
A test with $1,310,000$ irregularly spaced field data over the Caraj{\'a}s Province, Brazil, 
confirms the efficiency of our method by taking $\approx 385.56$ seconds to estimate the physical-property
distribution over the equivalent layer and $\approx 2.64$ seconds to compute the upward 
continuation.

@@

Large-scale data processing with tens of thousands of data is a reality in all areas of geophysics including the geophysical potential fields. 
The potential-field data processing includes convolution integrals which can be solved either in the space or Fourier domains.
The earliest techniques of potential-field data processing  were developed in the space domain.
For example, \cite{Peters1949} accomplished, in the space domain, the second and fourth derivatives of magnetic data and the upward- and downward-continuations of magnetic data by deriving coefficients that are
used in a  graphical convolution with the magnetic data.
However, the techniques for  processing potential-field data in space-domain were soon substituted by the Fourier-domain techniques. 
\cite{Dean1958} pointed out that the operations of second order derivatives, analytic continuation, smoothing, the removing of residuals or regionals, and others for processing  potential-field data are
similar to the electric filter circuits in Fourier domain.
\cite{Dean1958} was the first to develop the theory of linear filters in Fourier domain for gravity and magnetic processing and to present filters in Fourier domain \citep[][ see Table I, p 113]{Dean1958} 
for some theoretical geophysical operations (e.g., derivatives and upward and downward continuations).
\cite{GUNN1975} presented a comprehensive analysis of processing potential-field data in Fourier domain.

An approach for processing potential-field data in space domain is the equivalent-layer technique.
The equations deductions of the equivalent layer as a solution of the Laplace's equation in the region above the source was first presented by \cite{kellogg1929} and detailed explanations can also be found in \cite{blakely1996}. 
Although the equivalent-layer technique  has been known since the 1960s in geophysical literature \citep{danes1961structure,bott1967solution,dampney1969}, its use has become feasible only recently 
because of the advances in computational power.
In magnetic data procesing, some authors explored this technique for calculating the first and second vertical derivatives of the fields \citep{emilia1973}, reduction to the pole \citep{silva1986,oliveirajr-etal2013,li2014using}, upward/downward continuations \citep{hansen-miyazaki1984,li-oldenburg2010} and total magnetic induction vector components calculation \citep{sun2019constrained}.

Together with the rise in computational processing power, some works tried new implementations to increase the efficiency of the equivalent layer. 
In \cite{leao-silva1989} the authors used a shifting window over the layer to increase the number of linear systems to be solved, but to reduced the size of each linear system at the same time. 
Another approach for a fast equivalent layer was proposed by \cite{li-oldenburg2010}  who transformed the full sensitivity matrix into a sparse one by using  the compression of the coefficient matrix 
using fast wavelet transforms based on orthonormal, compactly supported wavelets.  
\cite{oliveirajr-etal2013} divided the equivalent layer into a grid of fixed source windows.
Instead of directly calculating the the physical-property distribution of a finite set of equivalent
sources (e.g., dipoles, point of masses) arranged in the entire equivalent layer,
\cite{oliveirajr-etal2013} estimated the coefficients of a bivariate polynomial function describing 
the physical-property distribution within each equivalent-source window.
The estimated polynomial-coefficients are transformed into the physical-property distribution
and thus any standard linear transformation of the data can be performed.
Grounded on excess mass constraint, \cite{siqueira-etal2017} proposed an iterative method 
for processing large gravimetric data using the equivalent layer without requiring the solution 
of a linear system. 
In \cite{siqueira-etal2017}, the initial mass distribution over the equivalent layer is
proportional to observed gravity data and it is updated at each iteration by adding mass corrections that are proportional to the residuals of observed and estimated data.

One of the greatest obstacles to the use of the equivalent-layer technique for processing potential-field data is the solution of the associated linear system.
A wide variety of applications in mathematics and engineering that fall into Toeplitz systems propelled the development of a large variety of  methods for solving them. Direct methods were conceived by \cite{levinson1946} and by \cite{trench1964}. Currently, the iterative method of conjugate gradient is used in most cases, in \cite{chan-jin2007} the authors presented an introduction on the topic for 1D data structures of Toeplitz matrices and also for 2D data structures, which they called block-Toeplitz Toeplitz-block matrices. In both cases, the solving strategy is to embed the Toeplitz/BTTB matrix into a Circulant/Block-Circulant Circulant-Block matrix, calculate its eigenvalues by a 1D or 2D fast Fourier transform of its first column, respectively and carry the matrix-vector product between kernel and parameters at each iteration of the conjugate gradient method in a very fast manner.

@@

In this work, the convolutional equivalent layer using the block-Toeplitz Toeplitz-block idea, presented in \cite{takahashi2020convolutional}, will be used to solve the linear system required to estimate the physical property that produces a magnetic field on regular grids. 
Here, we achieve very fast solutions using a conjugate gradient algorithm combined with the fast Fourier transform. We present a novel method of exploring the symmetric structures of the second derivatives of the inverse of the distance contained in the magnetic kernel, to keep the memory RAM usage to the minimal by using only one equivalent source to carry the calculations of the forward problem. We also show tests of the magnetic convolutional equivalent layer when irregular grids are used. The convergence of the conjugate gradient maintains in an acceptable level even using irregular grids. 
Our results show the good performance of our method in producing fast and robust solutions for processing large amounts of magnetic data using the equivalent layer technique.

@@

Let $\mathbf{d}^{o}$ be the $N \times 1$ observed data vector $(x_{i}, y_{i}, z_{i})$, $i =  1, \dots, N $,
whose $i$th element is the total-field anomaly $d^{o}_{i}$ produced by arbitrarily magnetized sources
aranged in a right-handed Cartesian coordinate system with $x$-, $y$- and $z$-axis 
pointing to north, east and down, respectively.
We consider that the total-field anomaly data $d^{o}_{i}$ represent the discrete
values of a harmonic function. Besides, we consider that the main geomagnetic field 
direction is constant at the limited study area and can be defined by the unit vector
\begin{equation}
	\hat{\mathbf{F}} = \begin{bmatrix}
		F_x \\
		F_y \\
		F_z
	\end{bmatrix} =
	\begin{bmatrix}
		\cos(I_{0}) \, \cos(D_{0}) \\
		\cos(I_{0}) \, \sin(D_{0}) \\
		\sin(I_{0})
	\end{bmatrix} \: ,
	\label{eq:unit_vector_F}
\end{equation}
with constant inclination $I_{0}$ and declination $D_{0}$.
In this case, $d^{o}_{i}$ can be approximated by the predicted total-field anomaly
\begin{equation}
	\Delta T_{i} = \sum_{j=1}^{M} \, p_{j} a_{ij} \: ,
	\label{eq:integral-sum_mag}
\end{equation}
which describes the magnetic induction exerted at the observation point,
by a discrete layer of $M$ dipoles (equivalent sources) defined on the horizontal plane $z = z_{c}$, 
where $p_{j}$ is the magnetic moment intensity (in A~m~$^{2}$)~of the $j$th dipole, 
that has unit volume and is located at the point $(x_{j}, y_{j}, z_{c})$. In equation
\ref{eq:integral-sum_mag}, $a_{ij}$ is the harmonic function
\begin{equation}
	a_{ij}
	= c_{m} \, \frac{\mu_{0}}{4\pi} \, \hat{\mathbf{F}}^{\top} \mathbf{H}_{ij} \: \hat{\mathbf{u}} \: ,
	\label{eq:aij_mag}
\end{equation}
the unit vector
\begin{equation}
	\hat{\mathbf{u}} = \begin{bmatrix}
		u_x \\
		u_y \\
		u_z
	\end{bmatrix} =
	\begin{bmatrix}
		\cos(I) \, \cos(D) \\
		\cos(I) \, \sin(D) \\
		\sin(I)
	\end{bmatrix} \: ,
	\label{eq:u_hat}
\end{equation}
defines the magnetization direction of all dipoles, with constant inclination $I$ and declination $D$,
$\mu_{0} = 4\pi \, 10^{-7}$ H/m is the magnetic constant, $c_{m} = 10^{9}$ is a factor that transforms
the magnetic induction from Tesla (T) to nanotesla (nT) and $\mathbf{H}_{ij}$ is a $3 \times 3$ matrix 
\begin{equation}
	\mathbf{H}_{ij} = \begin{bmatrix}
		h^{xx}_{ij} & h^{xy}_{ij} & h^{xz}_{ij} \\
		h^{xy}_{ij} & h^{yy}_{ij} & h^{yz}_{ij} \\
		h^{xz}_{ij} & h^{yz}_{ij} & h^{zz}_{ij}
	\end{bmatrix} \: ,
	\label{eq:Hij}
\end{equation}
where 
\begin{equation}
	h^{\alpha\beta}_{ij} = 
	\begin{cases}
		\frac{3 \left( \alpha_{i} - \alpha_{j} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: , \quad \alpha = \beta \\
		\frac{3 \left( \alpha_{i} - \alpha_{j} \right) \left( \beta_{i} - \beta_{j} \right)}{r_{ij}^{5}} \: , \quad \alpha \ne \beta
	\end{cases} \: , \quad \alpha, \beta = x, y, z \: ,
	\label{eq:hij_alpha_beta}
\end{equation}
are the second derivatives of the inverse distance function
\begin{equation}
	\frac{1}{r_{ij}} = 
	\frac{1}{\sqrt{\left(x_{i} - x_{j} \right)^{2} + 
			\left(y_{i} - y_{j} \right)^{2} + \left(z_{i} - z_{c} \right)^{2}}}
	\label{eq:1_rij}
\end{equation}
with respect to the coordinates of the observation point $(x_{i}, y_{i}, z_{i})$.

Equation \ref{eq:integral-sum_mag} can be rewritten in matrix notation as follows:
\begin{equation}
	\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
	\label{eq:predicted-data-vector_mag}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is the $N \times 1$ predicted data vector with $i$th element defined
as the predicted total-field anomaly $\Delta T_{i}$ (equation \ref{eq:integral-sum_mag}),
$\mathbf{p}$ is the $M \times 1$ parameter vector whose $j$th element is the magnetic moment intensity
$p_{j}$ of the $j$th dipole and $\mathbf{A}$ is the $N \times M$ sensitivity matrix with element 
$ij$ defined by the harmonic function $a_{ij}$ (equation \ref{eq:aij_mag}).
In the classical equivalent-layer technique, the common approach for 
estimating the parameter vector $\mathbf{p}$ from the observed 
total-field anomaly data $\mathbf{d}^{o}$ is solving the least-squares normal equations
\begin{equation}
	\mathbf{A}^{\top}\mathbf{A} \: \mathbf{p} = 
	\mathbf{A}^{\top} \mathbf{d}^{o} \: .
	\label{eq:normal-equations}
\end{equation}
Equation \ref{eq:normal-equations} is usually solved by first computing the Cholesky 
factor $\mathbf{G}$ of the positive-definite matrix $\mathbf{A}^{\top}\mathbf{A}$ and then using it to solve the linear 
systems \citep[][ p. 262]{golub-vanloan2013}:
\begin{equation}
	\begin{split}
		\mathbf{G} \mathbf{w} &= \mathbf{A}^{\top}\mathbf{d}^{o} \\
		\mathbf{G}^{\top} \tilde{\mathbf{p}} &= \mathbf{w}
	\end{split} \: ,
	\label{eq:classical-method}
\end{equation}
where $\mathbf{w}$ is a temporarily variable.
This approach to estimating the parameter vector will be 
referenced throughout this work as the \textit{classical method}.
The computational cost associated with the classical method can be very high
when dealing with large datasets. In the following subsections, we will show how to 
explore the structure of the sensitivity matrix $\mathbf{A}$ and 
efficiently solve the least-squares normal equations (equation \ref{eq:normal-equations}).

@@

Now we can define the structure of $\mathbf{A}$ in terms of its components 
$\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}). To do this, 
we consider the particular case in which the observed total-field anomaly is located 
on an $N_{x} \times N_{y}$ 
regular grid of points spaced by $\Delta_{x}$ and $\Delta_{y}$ along the $x$- and $y$-directions,
respectively, on a constant vertical coordinate $z_{0}$. We also consider that the equivalent layer
is aranged such that one dipole is located right below each observation point, at a constant coordinate $z_{c}$.
In this case, the number of equivalent sources $M$ is equal to the number of data $N$ and, 
consequently, matrices $\mathbf{A}$ and $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ become 
square ($N \times N$). 
Besides, the horizontal coordinates $x_{i}$ and $y_{i}$ of the observation points 
can be defined by
\begin{equation}
	x_{i} = x_{1} + \left[ k(i) - 1 \right] \, \Delta_{x}
	\label{eq:xi}
\end{equation}
and
\begin{equation}
	y_{i} = y_{1} + \left[ l(i) - 1 \right] \, \Delta_{y} \: ,
	\label{eq:yi}
\end{equation}
where $x_{1}$ and $y_{1}$ are the start grid coordinates for $x_{i}$ and $y_{i}$, respectively,
and $k(i)$ and $l(i)$ are integer functions defined according to the orientation
of the data grid (Figure \ref{fig:regular-grids}). 
For $x$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x}
	\label{eq:k-x-oriented}
\end{equation}
and
\begin{equation}
	l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil \: .
	\label{eq:l-x-oriented}
\end{equation}
For $y$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
	\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
	l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \: .
	\label{eq:l-y-oriented}
\end{equation}
In equations \ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}, $\lceil \cdot \rceil$ denotes the ceiling 
function \citep[e.g.,][ p. 67-68]{graham-etal1994}.
Equations \ref{eq:xi}--\ref{eq:l-y-oriented} can also be used to define the coordinates 
$x_{j}$ and $y_{j}$ of the equivalent sources, but with index $j$ instead of $i$.

>>>>

		We present a fast equivalent layer method for processing large-scale magnetic data. 
		We demonstrate that the sensitivity matrix associated with an equivalent layer
		of dipoles assumes a Block-Toeplitz Toeplitz-Block (BTTB) structure for the 
		case in which observations and dipoles are aligned on a 
		horizontal and regularly-spaced grid.
		The product of a BTTB matrix and an arbitrary vector represents a discrete 
		convolution and can be efficiently computed via 2D Fast Fourier Transform.
		In this case, the matrix-vector product uses only the elements forming the first column
		of the BTTB matrix, saving computational time and memory. 
		Our convolutional equivalent layer method uses this approach to compute 
		the matrix-vector products in the iterative conjugate gradient algorithm with the purpose 
		of estimating the physical-property distribution over the equivalent layer with 
		large data sets.
		Synthetic tests with a mid-size $100 \times 50$ grid of total-field anomaly data
		show a decrease of $\approx 10^4$ in floating-point operations and $\approx 25\times$ 
		in computation runtime of our method compared to the classical approach of solving
		the least-squares normal equations via Cholesky decomposition. 
		Better results are obtained for millions of data, showing drastic decreases in RAM usage
		and runtime, allowing to perform magnetic data processing of large data sets on regular 
		desktop computers. 
		Our results also show that, compared to the classical Fourier approach, the magnetic
		data processing with our method requires similar computation time, but produces significantly 
		smaller border effects without using any padding scheme and also is much more robust to 
		deal with data on irregularly spaced points or on undulating observation surfaces.
		A test with $1,310,000$ irregularly spaced field data over the Caraj{\'a}s Province, Brazil, 
		confirms the success of our method by taking $\approx 385.56$ seconds to estimate the physical-property
		distribution over the equivalent layer and $\approx 2.64$ seconds to compute the upward 
		continuation.

@@

	Large-scale data processing with tens of thousands of data, is a reality in all areas of geophysics including the geophysical potential fields. 
	The potential-field data processing includes convolution integrals which can be solved either in the space or Fourier domains.
	The earliest techniques of potential-field data processing  were developed in the space domain.
	For example, \cite{Peters1949} accomplished, in the space domain, the second and fourth derivatives of magnetic data and the upward- and downward-continuations of magnetic data by deriving coefficients that are
	used in a  graphical convolution with the magnetic data.
	Howerver, the techniques for  processing potential-field data in space-domain were soon substituted by the Fourier-domain techniques. 
	\cite{Dean1958} pointed out that the operations of second derivative, analytic continuation, smoothing, the removing of residuals or regionals, and others for processing  potential-field data are
	similar to the electric filter circuits in Fourier domain.
	\cite{Dean1958} was the first to develop the theroy of  linear filter in Fourier domain for gravity and magnetic processing and to present filters in Fourier domain \citep[][ see Table I, p 113]{Dean1958} 
	for some theoretical geophysical operations (e.g., derivatives and upward and downward continuations).
	\cite{GUNN1975} presented a comprehensive analysis of processing potential-field data in Fourier domain.

	An approach for processing potential-field data in space domain is the equivalent-layer technique.
	The equations deductions of the equivalent layer as a solution of the Laplace's equation in the region above the source was first presented by \cite{kellogg1929} and detailed explanations can also be found in \cite{blakely1996}. 
	Although the equivalent-layer technique  has been known since the 1960s in geophysical literature \citep{danes1961structure,bott1967solution,dampney1969}, its use has become feasible only recently 
	because the advances in computational power.
	In magnetic data procesing, some authors explored this technique for calculating the first and second vertical derivatives fields \citep{emilia1973}, reduction to the pole \citep{silva1986,oliveirajr-etal2013,li2014using}, upward/downward continuations \citep{hansen-miyazaki1984,li-oldenburg2010} and total magnetic induction vector components calculation \citep{sun2019constrained}.

	Together with the rise in computational processing power, some works tried new implementations to increase the efficiency of the equivalent layer. 
	In \cite{leao-silva1989} the authors used a shifting window over the layer, increasing the number of linear systems to be solved, but the size of each linear system is reduced. 
	Another approach for a fast equivalent layer was proposed by \cite{li-oldenburg2010}  who transformed the full sensitivity matrix into a sparse one by using  the compression of the coefficient matrix 
	using fast wavelet transforms based on orthonormal, compactly supported wavelets.  
	\cite{oliveirajr-etal2013} divided the equivalent layer into a grid of fixed source windows.
	Instead of directly calculating the the physical-property distribution of a finite set of equivalent
	sources (e.g., dipoles, point of masses) arranged in the entire equivalent layer,
	\cite{oliveirajr-etal2013} estimated the coefficients of a bivariate polynomial function describing 
	the physical-property distribution within each equivalent-source window.
	The estimated polynomial-coefficients are transformed into the physical-property distribution
	and thus any standard linear transformation of the data can be performed.
	Grounded on excess mass constraint, \cite{siqueira-etal2017} proposed an iterative method 
	for processing large gravimetric data using the equivalent layer without requiring the solution 
	of a linear system. 
	In \cite{siqueira-etal2017}, the initial mass distribution over the equivalent layer is
	proportional to observed gravity data and it is updated at each iteration by adding mass corrections that are proportional to the residuals of observed and estimated data.

	One of the greatest obstacles to the use of the equivalent-layer technique for processing potential-field data is the solution of the associated linear system.
	A wide variety of applications in mathematics and engineering that fall into Toeplitz systems propelled the development of a large variety of  methods for solving them. Direct methods were conceived by \cite{levinson1946} and by \cite{trench1964}. Currently, the iterative method of conjugate gradient is used in most cases, in \cite{chan-jin2007} the authors presented an introduction on the topic for 1D data structures of Toeplitz matrices and also for 2D data structures, which they called block-Toeplitz Toeplitz-block matrices. In both cases, the solving strategy is to embbed the Toeplitz/BTTB matrix into a Circulant/Block-Circulant Circulant-Block matrix, calculate its eigenvalues by a 1D or 2D fast Fourier transform of its first column, respectively and carry the matrix-vector product between kernel and parameters at each iteration of the conjugate gradient method in a very fast manner.

@@

	In this work, the convolutional equivalent layer using the block-Toeplitz Toeplitz-block idea, presented in \cite{takahashi2020convolutional}, will be used to solve the linear system required to estimate the physical property that produces a magnetic field on regular grids. 
	Here, we achive very fast solutions using a conjugate gradient algorithm combined with the fast Fourier transform. We present a novel method of exploring the symmetric structures of the second derivatives of the inverse of the distance contained in the magnetic kernel, to keep the memory RAM usage to the minimal by using only one equivalent source to carry the calculations of the forward problem. We also show tests of the magnetic convolutional equivalent layer when irregular grids are used. The convergence of the conjugate gradient maintains in an acceptable level even using irregular grids. 
	Our results show the good performance of our method in producing fast and robust solutions for processing large amounts of magnetic data using the equivalent layer technique.

@@

	Let $\mathbf{d}^{o}$ be the $N \times 1$ observed data vector, whose $i$th element 
	is the total-field anomaly $d^{o}_{i}$ produced by arbitrarily magnetized sources
	at the position $(x_{i}, y_{i}, z_{i})$, $i =  1, \dots, N $, 
	of a right-handed Cartesian coordinate system with $x$-, $y$- and $z$-axis 
	pointing to north, east and down, respectively.
	We consider that the total-field anomaly data $d^{o}_{i}$ represent the discrete
	values of a harmonic function. Besides, we consider that the main geomagnetic field 
	direction at the study area can be defined by the unit vector
	\begin{equation}
		\hat{\mathbf{F}} = \begin{bmatrix}
			F_x \\
			F_y \\
			F_z
		\end{bmatrix} =
		\begin{bmatrix}
			\cos(I_{0}) \, \cos(D_{0}) \\
			\cos(I_{0}) \, \sin(D_{0}) \\
			\sin(I_{0})
		\end{bmatrix} \: ,
		\label{eq:unit_vector_F}
	\end{equation}
	with constant inclination $I_{0}$ and declination $D_{0}$.
	In this case, $d^{o}_{i}$ can be approximated by the predicted total-field anomaly
	\begin{equation}
		\Delta T_{i} = \sum_{j=1}^{M} \, p_{j} a_{ij} \: ,
		\label{eq:integral-sum_mag}
	\end{equation}
	which describes the magnetic induction exerted, at the observation point $(x_{i}, y_{i}, z_{i})$,
	by a discrete layer of $M$ dipoles (equivalent sources) defined on the horizontal plane $z = z_{c}$, 
	where $p_{j}$ is the magnetic moment intensity (in A~m~$^{2}$)~of the $j$th dipole, 
	that has unit volume and is located at the point $(x_{j}, y_{j}, z_{c})$. In equation
	\ref{eq:integral-sum_mag}, $a_{ij}$ is the harmonic function
	\begin{equation}
		a_{ij}
		= c_{m} \, \frac{\mu_{0}}{4\pi} \, \hat{\mathbf{F}}^{\top} \mathbf{H}_{ij} \: \hat{\mathbf{u}} \: ,
		\label{eq:aij_mag}
	\end{equation}
	the unit vector
	\begin{equation}
		\hat{\mathbf{u}} = \begin{bmatrix}
			u_x \\
			u_y \\
			u_z
		\end{bmatrix} =
		\begin{bmatrix}
			\cos(I) \, \cos(D) \\
			\cos(I) \, \sin(D) \\
			\sin(I)
		\end{bmatrix} \: ,
		\label{eq:u_hat}
	\end{equation}
	defines the magnetization direction of all dipoles, with constant inclination $I$ and declination $D$,
	$\mu_{0} = 4\pi \, 10^{-7}$ H/m is the magnetic constant, $c_{m} = 10^{9}$ is a factor that transforms
	the magnetic induction from Tesla (T) to nanotesla (nT) and $\mathbf{H}_{ij}$ is a $3 \times 3$ matrix 
	\begin{equation}
		\mathbf{H}_{ij} = \begin{bmatrix}
			h^{xx}_{ij} & h^{xy}_{ij} & h^{xz}_{ij} \\
			h^{xy}_{ij} & h^{yy}_{ij} & h^{yz}_{ij} \\
			h^{xz}_{ij} & h^{yz}_{ij} & h^{zz}_{ij}
		\end{bmatrix} \: ,
		\label{eq:Hij}
	\end{equation}
	where 
	\begin{equation}
		h^{\alpha\beta}_{ij} = 
		\begin{cases}
			\frac{3 \left( \alpha_{i} - \alpha_{j} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: , \quad \alpha = \beta \\
			\frac{3 \left( \alpha_{i} - \alpha_{j} \right) \left( \beta_{i} - \beta_{j} \right)}{r_{ij}^{5}} \: , \quad \alpha \ne \beta
		\end{cases} \: , \quad \alpha, \beta = x, y, z \: ,
		\label{eq:hij_alpha_beta}
	\end{equation}
	are the second derivatives of the inverse distance function
	\begin{equation}
		\frac{1}{r_{ij}} = 
		\frac{1}{\sqrt{\left(x_{i} - x_{j} \right)^{2} + 
				\left(y_{i} - y_{j} \right)^{2} + \left(z_{i} - z_{c} \right)^{2}}}
		\label{eq:1_rij}
	\end{equation}
	with respect to the coordinates of the observation point $(x_{i}, y_{i}, z_{i})$.

	Equation \ref{eq:integral-sum_mag} can be rewritten in matrix notation as follows:
	\begin{equation}
		\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
		\label{eq:predicted-data-vector_mag}
	\end{equation}
	where $\mathbf{d}(\mathbf{p})$ is the $N \times 1$ predicted data vector with $i$th element defined
	as the predicted total-field anomaly $\Delta T_{i}$ (equation \ref{eq:integral-sum_mag}),
	$\mathbf{p}$ is the $M \times 1$ parameter vector whose $j$th element is the magnetic moment intensity
	$p_{j}$ of the $j$th dipole and $\mathbf{A}$ is the $N \times M$ sensitivity matrix with element 
	$ij$ defined by the harmonic function $a_{ij}$ (equation \ref{eq:aij_mag}).
	In the classical equivalent-layer technique, the common approach for 
	estimating the parameter vector $\mathbf{p}$ from the observed 
	total-field anomaly data $\mathbf{d}^{o}$ is solving the least-squares normal equations
	\begin{equation}
		\mathbf{A}^{\top}\mathbf{A} \: \mathbf{p} = 
		\mathbf{A}^{\top} \mathbf{d}^{o} \: .
		\label{eq:normal-equations}
	\end{equation}
	Equation \ref{eq:normal-equations} is usually solved by first computing the Cholesky 
	factor $\mathbf{G}$ of matrix $\mathbf{A}^{\top}\mathbf{A}$ and then using it to solve the linear 
	systems \citep[][ p. 262]{golub-vanloan2013}:
	\begin{equation}
		\begin{split}
			\mathbf{G} \mathbf{w} &= \mathbf{A}^{\top}\mathbf{d}^{o} \\
			\mathbf{G}^{\top} \tilde{\mathbf{p}} &= \mathbf{w}
		\end{split} \: ,
		\label{eq:classical-method}
	\end{equation}
	where $\mathbf{w}$ is a dummy variable.
	This approach to estimate the parameter vector will be 
	referenced throughout this work as the \textit{classical method}.
	The computational cost associated with the classical method can be very high
	when dealing with large datasets. In the following subsections, we will show how to 
	explore the structure of the sensitivity matrix $\mathbf{A}$ and 
	efficiently solve the least-squares normal equations (equation \ref{eq:normal-equations}).

@@

	Now we can define the structure of $\mathbf{A}$ in terms of its components 
	$\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}). To do this, 
	we consider the particular case in which the observed total-field anomaly is located 
	on an $N_{x} \times N_{y}$ 
	regular grid of points spaced by $\Delta_{x}$ and $\Delta_{y}$ along the $x$- and $y$-directions,
	respectively, on a constant vertical coordinate $z_{0}$. We also consider that the equivalent layer
	is formed by one dipole right below each observation point, at the constant coordinate $z_{c}$.
	In this case, the number of equivalent sources $M$ is equal to the number of data $N$ and, 
	consequently, matrices $\mathbf{A}$ and $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ become 
	square ($N \times N$). 
	Besides, the horizontal coordinates $x_{i}$ and $y_{i}$ of the observation points 
	can be defined by
	\begin{equation}
		x_{i} = x_{1} + \left[ k(i) - 1 \right] \, \Delta_{x}
		\label{eq:xi}
	\end{equation}
	and
	\begin{equation}
		y_{i} = y_{1} + \left[ l(i) - 1 \right] \, \Delta_{y} \: ,
		\label{eq:yi}
	\end{equation}
	where $x_{1}$ and $y_{1}$ are the lower limits for $x_{i}$ and $y_{i}$, respectively,
	and $k(i)$ and $l(i)$ are integer functions defined according to the orientation
	of the data grid (Figure \ref{fig:regular-grids}). 
	For $x$-\textit{oriented grids}, the integer functions are given by
	\begin{equation}
		k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x}
		\label{eq:k-x-oriented}
	\end{equation}
	and
	\begin{equation}
		l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil \: .
		\label{eq:l-x-oriented}
	\end{equation}
	For $y$-\textit{oriented grids}, the integer functions are given by
	\begin{equation}
		k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
		\label{eq:k-y-oriented}
	\end{equation}
	and
	\begin{equation}
		l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \: .
		\label{eq:l-y-oriented}
	\end{equation}
	In equations \ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}, $\lceil \cdot \rceil$ denotes the ceiling 
	function \citep[e.g.,][ p. 67-68]{graham-etal1994}.
	Equations \ref{eq:xi}--\ref{eq:l-y-oriented} can also be used to define the coordinates 
	$x_{j}$ and $y_{j}$ of the equivalent sources, but with index $j$ instead of $i$.

	By using equations \ref{eq:xi}--\ref{eq:l-y-oriented} to define the coordinates $x_{i}$ and 
	$y_{i}$ of the observation points and $x_{j}$ and $y_{j}$ of the equivalent sources, we can
	rewrite the elements $h^{\alpha\beta}_{ij}$ (equation \ref{eq:hij_alpha_beta}) of matrix 
	$\mathbf{H}_{ij}$ (equation \ref{eq:Hij}) as follows:
	\begin{equation}
		h^{xx}_{ij} = 
		\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
		\label{eq:hxx_regular}
	\end{equation}
	\begin{equation}
		h^{yy}_{ij} = 
		\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
		\label{eq:hyy_regular}
	\end{equation}
	\begin{equation}
		h^{zz}_{ij} = 
		\frac{3 \Delta_{z}^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
		\label{eq:hzz_regular}
	\end{equation}
	\begin{equation}
		h^{xy}_{ij} = 
		\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)\left( \Delta l_{ij} \, \Delta_{y} \right)}{r_{ij}^{5}} \: ,
		\label{eq:hxy_regular}
	\end{equation}
	\begin{equation}
		h^{xz}_{ij} = 
		\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right) \Delta_{z}}{r_{ij}^{5}}
		\label{eq:hxz_regular}
	\end{equation}
	and
	\begin{equation}
		h^{yz}_{ij} = 
		\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right) \Delta_{z}}{r_{ij}^{5}} \: ,
		\label{eq:hyz_regular}
	\end{equation}
	where $\Delta_{z} = z_{c} - z_{0}$, 
	\begin{equation}
		\Delta k_{ij} = \frac{x_{i} - x_{j}}{\Delta_{x}} = k(i) - k(j) \: ,
		\label{eq:Delta_kij}
	\end{equation}
	\begin{equation}
		\Delta l_{ij} = \frac{y_{i} - y_{j}}{\Delta_{y}} = l(i) - l(j)
		\label{eq:Delta_lij}
	\end{equation}
	and 
	\begin{equation}
		\frac{1}{r_{ij}} = 
		\frac{1}{\sqrt{\left( \Delta k_{ij} \, \Delta_{x} \right)^{2} + \left( \Delta l_{ij} \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
		\label{eq:1_rij_regular}
	\end{equation}
	Note that the integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ (equations 
	\ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}) defining $\Delta k_{ij}$ (equation
	\ref{eq:Delta_kij}), $\Delta l_{ij}$ (equation \ref{eq:Delta_lij}) and 
	$\tfrac{1}{r_{ij}}$ (equation \ref{eq:1_rij_regular}) assume different 
	forms depending on the grid orientation.
	Despite of that, it can be shown that
	\begin{equation}
		\Delta k_{ij} = - \Delta k_{ji} \: ,
		\label{eq:Delta_kij_symmetry}
	\end{equation}
	\begin{equation}
		\Delta l_{ij} = - \Delta l_{ji}
		\label{eq:Delta_lij_symmetry}
	\end{equation}
	and 
	\begin{equation}
		\frac{1}{r_{ij}} = \frac{1}{r_{ji}}
		\label{eq:1_rij_symmetry}
	\end{equation}
	for any grid orientation.
